\cursedforest\ - A Random Forest Implementation for ``Big'' and ``Wide'' Data \\ Supplementary Information

### 2017-01-13 doc/publications/plos16/plos16.tex
some more changes

%% Additional packages
\usepackage{footmisc}
\usepackage{multirow}
\usepackage[T1]{fontenc}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Section 2: Feature Selection on real data 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\cursedforest\ selects previously demonstrated bone-density variants}
We consider the problem of feature discovery on a genomic dataset for a specific phenotype.
In this case we investigate the ability of  ability to locate variants in the same loci as genes previously demonstrated to
be associated with a high or low bone mass density (BMD) and rank these variants as being "important".
From a random forest model built on this data, we produce a list of the to 5,000 features. 
Applying the Mann–Whitney U test to these fetures we observe a p-value of x, supporting our
hypothesis that features in close proximity to the known BMD genes are Contributing more information to our classification model.

One observation we make is that of the most-important features---for example, the top ten variants ranked by Gini impurity---is
that not all of these are variants are in close proximity of known BMD genes.
One of these is COLEC10. Featuring a transcription factor binding site for FOXL1, COLEC10 is also reported as containing a SNP
associated with BMD \cite{Liu2008}.

@@

@@

We note several things here:
\begin{itemize}
\item the Donoho-Tanner phase transition (in this setting) refers to estimating the $\beta$ in data generated by a
  linear model;
\item in a decision tree (random forest) there is no notion of estimating the $\beta$
\item also a decision tree (random forest) is a heruistic search. It may recover a relationshipin the space
  of the combinatorial search.
\end{itemize}

@@

\begin{figure}[tbhp] 
 \begin{adjustwidth}{-1.00in}{0in}
    \centering
    \includegraphics[totalheight=6cm]{./figs/phase.png} 
    \caption{{\bf The phase chage diagram.}}
    \label{figure:phase-diagram-equivalence.png} 
    \vspace{4ex}
  \end{adjustwidth}
\end{figure}

>>>>

%% Additional packages
\usepackage{footmisc}
\usepackage{multirow}
%\usepackage[T1]{fontenc}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Section 2: Feature Selection on real data 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\cursedforest\ selects previously demonstrated bone-density variants}
We consider the problem of feature discovery on a genomic dataset for a specific phenotype.
In this case we investigate the ability to retrieve variants in the same loci as genes previously demonstrated to
be associated with a high or low bone mass density (BMD) and rank these variants as being "important".
We build a random forest model on this data, and produce a list of the top 5,000 contributing features, ranked by their
Gini impurity. We apply the Mann–Whitney U test to these fetures and observe a p-value of x, supporting our hypothesis
that features in close proximity to the known BMD genes are contributing more information to our classification model.

However, we also observe that from the set of highest-ranked features---for example, the top ten---not all of them exist
within close proximity to known BMD genes.

One example is COLEC10. Featuring a transcription factor binding site for FOXL1, COLEC10 is also reported as containing a SNP
associated with BMD \cite{Liu2008}.

We also frequently see highly-ranked variants in chromosome 22q11

@@

\marginpar{signal rather than ``significant variables''}

@@

We note several things here:
\begin{itemize}
\item 
\end{itemize}

@@

\begin{figure}[tbhp] 
 \begin{adjustwidth}{-1.00in}{0in}
    \centering
    \includegraphics[totalheight=6cm]{./figs/phase-diagram-equivalence.png} 
    \caption{{\bf The phase chage diagram.}}
    \label{figure:phase-diagram-equivalence.png} 
    \vspace{4ex}
  \end{adjustwidth}
\end{figure}

\section{not for the paper}
Notes:
\begin{itemize}
\item we try a small simulation, a regression problem with $\beta(1:k) \sim U(1,100)$ and $y= X\beta + \epsilon$, $\epsilon \sim  N(0,\sqrt{16})$
\item \cite{Donoho.and.Stodden.2006} evaluate variable selection by 
$$\frac{\Vert\hat{\beta}-\beta\Vert_2}{\Vert\beta\Vert_2}$$
See figure \ref{figure:stepwise_lars_cv.png} 
\item 
However there is no $\hat{\beta}$ in a tree model. We are using variable importance as
a measure of the ordering of $\hat{\beta}$ (i.e. we expect variable importance to be proportional to $\hat{\beta}$). Is this correct?
\item as RF is sort of greedy approximation to combinatorial approach, it may have some success in the dead zone
\item  \cite{Donoho.and.Stodden.2006} are considering problems with $L_2$ errors (with additional  $L_1$ norms on the
  $\beta$. $L_1$).
How will their results carry over to 0-1 loss?
\end{itemize}

\begin{figure}[tbhp] 
  \begin{adjustwidth}{-1.00in}{0in}
    \begin{subfigure}[b]{0.5\linewidth}
      \centering
      \includegraphics[totalheight=6cm]{./figs/k.png}
      \caption{k}
      \label{figure:k.png}
      \vspace{4ex}
    \end{subfigure} 
    \begin{subfigure}[b]{0.5\linewidth}
      \centering
      \includegraphics[totalheight=6cm]{./figs/n.png}
      \caption{n}
      \label{figure:n.png}
    \end{subfigure} 
    \caption{}
  \end{adjustwidth}
\end{figure}

\begin{figure}[tbhp] 
  \begin{adjustwidth}{-1.00in}{0in}
    \begin{subfigure}[t]{0.5\linewidth}
      \centering
      \includegraphics[totalheight=6cm]{./figs/stepwise_lars_cv.png}
      \caption{Stepwise lars. $p=200$, $n$ and $k$ variable. Model selected by cross-validation}
      \label{figure:stepwise_lars_cv.png}
      \vspace{4ex}
    \end{subfigure} 
    \begin{subfigure}[t]{0.5\linewidth}
      \centering
      \includegraphics[totalheight=6cm]{./figs/ranger.png}
      \caption{ranger, RBO}
      \label{figure:ranger.png}
    \end{subfigure} 
    \caption{}
  \end{adjustwidth}
\end{figure}
